Direction-aware functional class scoring enrichment analysis of Infinium DNA methylation data

### 2024-02-14 manuscript/manuscript.Rmd
msedits
ms

![**Figure 8. Differential pathway methylation associated with prevalence of 14 common disease states.**
(A) A barplot showing the number of statistically significant pathways with higher and lower methylation identified in each prevalent condition (FDR<0.05).
(B) A heatmap of S scores for selected pathways across 14 common prevalent disease states.
Stars indicate that the pathway was identified as being among the top five differentially methylated pathways in each direction for each condition.
](img/fig8.svg)

@@

![**Figure 9. Differential pathway methylation associated with incidence of 19 common disease states.**
(A) A barplot showing the number of statistically significant pathways with higher and lower methylation identified in each incident condition (FDR<0.05).
(B) A heatmap of S scores for selected pathways across 19 common incident disease states.
Stars indicate that the pathway was identified as being among the top three differentially methylated pathways in each direction for each condition.
](img/fig9.svg)

@@

@@

This research was supported by use of the Nectar Research Cloud, a collaborative Australian research platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).
Authors gratefully acknowledge the contribution to this work of the Victorian Operational Infrastructure Support Program received by the Burnet Institute.
We thank the researchers who publicly shared data to enable this research.

>>>>

![**Figure 8. Differential pathway methylation associated with prevalence of 14 common disease states.**
(A) A bar plot showing the number of statistically significant pathways with higher and lower methylation identified in each prevalent condition (FDR<0.05).
(B) A heatmap of S scores for selected pathways across 14 common prevalent disease states.
Stars indicate that the pathway was identified as being among the top five differentially methylated pathways in each direction for each condition.
](img/fig8.svg)

@@

![**Figure 9. Differential pathway methylation associated with incidence of 19 common disease states.**
(A) A bar plot showing the number of statistically significant pathways with higher and lower methylation identified in each incident condition (FDR<0.05).
(B) A heatmap of S scores for selected pathways across 19 common incident disease states.
Stars indicate that the pathway was identified as being among the top three differentially methylated pathways in each direction for each condition.
](img/fig9.svg)

@@

## Data availability

Data sets reanalysed here are publicly available from NCBI GEO, Zenodo or from the supplementary tables of journal articles as described in the Methods section.

@@

This research was supported by use of the Nectar Research Cloud, a collaborative Australian research platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).
Authors gratefully acknowledge the contribution to this work of the Victorian Operational Infrastructure Support Program received by the Burnet Institute.
We thank the researchers who publicly shared data to enable this work.
